---
title: "Simulation Analysis with BayesSPM to check for Power and False Discovery Rate"
author: "Linda Amoafo"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
bibliography: reference.bib
vignette: >
  %\VignetteIndexEntry{Simulation Analysis with BayesSPM to check for Power and False Discovery Rate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

This is a second vignette run some simulations for the BayesSPM using the conservative threshold and the less conservative threshold. False Dicsovery Rates and Power are compared here. 

# Simulation Design 

Simulate time series of 10 time samples each. The smooth unit Gaussian random trajectories but with bursts at t = 5 and t = 7.

```{r}
set.seed(1)
grp=rep(c(0,1,2),each=20)
grp.matrix <- model.matrix(~factor(grp))
True_beta <- c(0.2,1,0.5)
time=1:10

Y <- matrix(NA, nrow=60, ncol=length(time))
for(i in 1:length(time)){
  if (i==5|i ==7){
    Y[,i] = rnorm(60, mean=grp.matrix %*% True_beta, sd=0.3)
  }else{Y[,i] = rnorm(60, mean=0, sd=0.3)}
}

colnames(Y) <- paste("Time", 1:10)

dat <- data.frame(grp, Y) |> tidyr::pivot_longer(cols=Time.1:Time.10) 
dat$name <- factor(dat$name, levels=paste("Time.", 1:10, sep=""), labels=paste("Time", 1:10))


ggplot2::ggplot(dat, ggplot2::aes(x=value))+
  ggplot2::geom_histogram(ggplot2::aes(y=ggplot2::after_stat(density),  fill=factor(grp)), alpha=0.5)+
  ggplot2::geom_density(ggplot2::aes(y=ggplot2::after_stat(density),  col=factor(grp)), alpha=.2) +
  ggplot2::facet_wrap(.~name, scale="free")+
  ggplot2::theme_bw()


graphics::matplot(t(Y), type = "l", col = matrix(grp+1, nrow=1))


```


## Simulations function
```{r}

Sim.Func <- function(i, True_beta=NULL){
   grp=rep(c(0,1,2),each=20)
   grp.matrix <- model.matrix(~factor(grp))
   if(is.null(True_beta)){True_beta <- c(0.5,1.5,2.5)}
   time=1:10
   
   Y <- matrix(NA, nrow=60, ncol=length(time))
   for(i in 1:length(time)){
  if (i==5|i ==7){
    Y[,i] = rnorm(60, mean=grp.matrix %*% True_beta, sd=0.5)
  }else{Y[,i] = rnorm(60, mean=0, sd=0.5)}
   }
   out <- SPMBayes(Y,group=grp, time, threshold=0.95, Q.threshold=0.05, Hypothesis="alt")
   
   list(PPH1=summary(out)$PPH1, QH1=summary(out)$QH1)
}



```


Output for this simulation was obtained from the high performance computer using the slurmR package.
The full path of the slurmfile.R and slurmfile.slurm files, which is included in the R package under:
inst/slurm_scripts/slurmfile.R and inst/slurm_scripts/slurmfile.slurm respectively. 

Here is how to access path for you,

```{r}
fn.slurm <- system.file(
  file.path("slurm_scripts", "slurmfile.slurm"),
  package = "BayesSPM"
  )

# fn.slurm now has the FULL PATH to the myfile.slurm file.
fn.slurm


fn.R <- system.file(
  file.path("slurm_scripts", "slurmfile.R"),
  package = "BayesSPM"
  )

# fn.slurm now has the FULL PATH to the myfile.slurm file.
fn.R
```

# Simulation Output Results

```{r}
SimBayesSPM <- readRDS("C:/Users/u1271266/OneDrive - University of Utah/Books/Spring 2023/PHS 7045/SPMR-package/BayesSPM/SimBayesSPM.rds")

PPH1.dat <- data.table::data.table(do.call(rbind,lapply(SimBayesSPM, "[[", "PPH1"))) |> 
  dplyr::mutate_all(.funs= \(x) x > 0.95) 

PPH1 <-   PPH1.dat |> dplyr::summarize_all(mean)  |> 
  tidyr::pivot_longer(cols=V1:V10, names_to = "Time", values_to = "Power") |> 
  dplyr::mutate(Time=as.numeric(factor(Time, levels=c("V1","V2","V3","V4","V5","V6","V7","V8","V9","V10"), labels=1:10)),Threshold="Conservative")

QH1.dat <- data.table::data.table(do.call(rbind,lapply(SimBayesSPM, "[[", "QH1"))) |> 
  dplyr::mutate_all(.funs= \(x) x < 0.05) 

QH1 <- QH1.dat |> dplyr::summarize_all(mean) |> 
  tidyr::pivot_longer(cols=V1:V10, names_to = "Time", values_to = "Power") |> 
  dplyr::mutate(Time=as.numeric(factor(Time, levels=c("V1","V2","V3","V4","V5","V6","V7","V8","V9","V10"), labels=1:10)), Threshold="Less Conservative")


ggplot2::ggplot(dplyr::bind_rows(PPH1, QH1), ggplot2::aes(y=Power, x=Time, color=Threshold)) +
  ggplot2::geom_line()+
  ggplot2::geom_hline(yintercept=0.8, linetype="dashed") +
  ggplot2::annotate("text", x = 4.5, y = 0.9, label =  paste("Conservative FDR = ", round(mean(as.matrix(PPH1.dat[,-c("V2","V7")])),3) , sep=""))+
  ggplot2::annotate("text", x = 4.5, y = 0.95, label =  paste("Less Conservative FDR = ", round(mean(as.matrix(PPH1.dat[,-c("V2","V7")])), 3) , sep="")) +
  ggplot2::theme_bw()

```








